---
title: 'Lipidomics analysis'
author: "Fanning Xia"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Purpose

To characterize and compare protein and lipid profiles from surface fish and 2 cavefish 2-cell stage embryos to see if there's any deposition difference as maternal nutrients.

# 0. Preparation

```{r}
# read the file from Compound Discover output
# filtering information see the PAR in proteomics drive 
# /n/proteomics/rohner/fx2482/PROT-922/Rohner-Xia_PROT-922-MAR_2023-12-14

# need to filter out the lipids that have average(sysH2O) or average(X-Bf) > 5 * min(ave(SF, PA and MO))


# set working directory
setwd("/Volumes/projects/fx2482/lipidomics-2cellstage/")

library(ggplot2)
library(NormalyzerDE)
library(preprocessCore)
library(dplyr)
library(limma)
library(tidyr)
library(ggfortify)
library(cluster)
library(ggsci)
library(devtools)
#install_github("metabolomicsworkbench/RefMet")
library(RefMet)
library(ggrepel)
library(tidyverse)
library(scales)
#show_col(pal_npg("nrc")(9))
library(gridExtra)
library(vangogh)
library(ggpubr)
library(ggExtra)
library(car)
library(FSA)
library(reshape2)


pvg35 <- c(vangogh_palette("StarryRhone"),
           vangogh_palette("SelfPortrait"),
           vangogh_palette("Irises"),
           vangogh_palette("CafeDeNuit"),
           vangogh_palette("Landscape"),
           vangogh_palette("Shoes"),
           vangogh_palette("SunflowersLondon"))
pvg20 <- c(vangogh_palette("Landscape"),
           vangogh_palette("Shoes"),
           vangogh_palette("CafeTerrace"), 
           vangogh_palette("Chaise"))
pvg15 <- pvg20[c(1:8,10,13:18)]
pvg3 <- c("#979FA2","#EE9B00","#93D2DF")
```


# 1. merge different rows (manually) and filter out some low quality lipid species

## 1.1 filter out low quality lipid species

#### positive ion mode

```{r}

# positive ion mode first

pos <- read.csv("input/individual/need_filtering_Am-Embryo_pos_6_compounds.csv")

pos$MO <- rowMeans(pos[, grep("^Group.Area..M", colnames(pos), value = TRUE)], na.rm = TRUE)
pos$PA <- rowMeans(pos[, grep("^Group.Area..P", colnames(pos), value = TRUE)], na.rm = TRUE)
pos$SF <- rowMeans(pos[, grep("^Group.Area..S\\d+", colnames(pos), value = TRUE)], na.rm = TRUE)


pos$SysH2O <- rowMeans(pos[, grep("^Group.Area..SysH2O", colnames(pos), value = TRUE)], na.rm = TRUE)
pos$Bf <- rowMeans(pos[, grep("^Group.Area..X.Bf", colnames(pos), value = TRUE)], na.rm = TRUE)

# keep every row if both mean(sysH2O) and mean(X-Bf) are lower than 5* min(mean(SF), mean(PA) and mean(MO))

pos_filtered <- pos[pos$SysH2O < 5*pos$MO &
                      pos$SysH2O < 5*pos$PA &
                      pos$SysH2O < 5*pos$SF &
                      pos$Bf < 5*pos$MO &
                      pos$Bf < 5*pos$PA &
                      pos$Bf < 5*pos$SF, ]


#write.csv(pos_filtered, "input/individual/negative control filtered_Am-Embryo_pos_6_compounds.csv")

rm(pos)
```

#### negative ion mode

```{r}
# negative ion mode then

neg <- read.csv("input/individual/need_filtering_Am-Embryo_neg_3_compounds.csv")

neg$MO <- rowMeans(neg[, grep("^Group.Area..M", colnames(neg), value = TRUE)], na.rm = TRUE)
neg$PA <- rowMeans(neg[, grep("^Group.Area..P", colnames(neg), value = TRUE)], na.rm = TRUE)
neg$SF <- rowMeans(neg[, grep("^Group.Area..S\\d+", colnames(neg), value = TRUE)], na.rm = TRUE)


neg$SysH2O <- rowMeans(neg[, grep("^Group.Area..SysH2O", colnames(neg), value = TRUE)], na.rm = TRUE)
neg$Bf <- rowMeans(neg[, grep("^Group.Area..X.Bf", colnames(neg), value = TRUE)], na.rm = TRUE)

# keep every row if both mean(sysH2O) and mean(X-Bf) are lower than 5* min(mean(SF), mean(PA) and mean(MO))

neg_filtered <- neg[neg$SysH2O < 5*neg$MO &
                      neg$SysH2O < 5*neg$PA &
                      neg$SysH2O < 5*neg$SF &
                      neg$Bf < 5*neg$MO &
                      neg$Bf < 5*neg$PA &
                      neg$Bf < 5*neg$SF, ]

#write.csv(neg_filtered, "input/individual/negative control filtered_Am-Embryo_neg_3_compounds.csv")

rm(neg)
```

## 1.2 merge the repetitive rows

```{r}

library(dplyr)
merged.pos <- pos_filtered %>%
  group_by(Name) %>%
  summarise(across(.cols = starts_with("Group.Area.."), .fns = sum))

merged.neg <- neg_filtered %>%
  group_by(Name) %>%
  summarise(across(.cols = starts_with("Group.Area.."), .fns = sum))

head(merged.pos)
head(merged.neg)

# check if row number is correct
length(unique(pos_filtered$Name))
length(unique(neg_filtered$Name))

rm(pos_filtered, neg_filtered)
```

## 1.3 check for pre-normalization matrix

#### used to decide which sample to keep

positive ion mode

```{r}
cols<-rep(c("darkcyan","darkorange3"),each=3)
boxplot(log2(merged.pos[,2:31]),col=cols,main="before norm in positive ion mode")
colSums(merged.pos[,2:31])
```

```{r}
cor_plot <- pheatmap::pheatmap(cor(log2(merged.pos[,2:31])), main = "before norm in positive ion mode")
#ggsave("output/vignette_run_outliersremoved_pos/median-norm/plots/before-norm-correlation-plot.png", plot = cor_plot,width = 10, height = 10, units = "in")
```

negative ion mode

```{r}
boxplot(log2(merged.neg[,2:24]),col=cols,main="before norm in negative ion mode")
colSums(merged.neg[,2:24])
```

```{r}
cor_plot <- pheatmap::pheatmap(cor(log2(merged.neg[,2:24])), main = "before norm in negative ion mode")
#ggsave("output/vignette_run_outliersremoved_neg/median-norm/plots/before-norm-correlation-plot.png", plot = cor_plot,width = 10, height = 10, units = "in")

rm(cor_plot)
```

decided to remove: M5 (N), P4(P), S3(P) and S1(N+P). Both positive and negative ion mode will be removed.

S3 positive ion mode cluster differently than all other surface replicates/samples. I've tried it before. S3 is still the outlier if we include it.

#### remove outliers and technical negative controls (water & buffer)

```{r}
merged.pos.filtered <- merged.pos[,c(1:5, 7:10, 12:13,15,17:19)]
boxplot(log2(merged.pos.filtered[,2:15]),col=cols,main="before norm in positive ion mode")
pheatmap::pheatmap(cor(log2(merged.pos.filtered[,2:15])), main = "before norm in positive ion mode")


merged.neg.filtered <- merged.neg[,c(1:9, 11:12,14,16:18)]
boxplot(log2(merged.neg.filtered[,2:15]),col=cols,main="before norm in negative ion mode")
pheatmap::pheatmap(cor(log2(merged.neg.filtered[,2:15])), main = "before norm in negative ion mode")

```

# 2. Normalization methods

## 2.1 load (and create) data and sampleinfo for heatmap

```{r}
#write.table(merged.pos.filtered[,2:15], file='input/individual/prep_normDE_outliersremoved_Am-Embryo_pos_6_compounds.tsv', quote=FALSE, sep='\t', col.names = NA)
#write.table(merged.neg.filtered[,2:15], file='input/individual/prep_normDE_outliersremoved_Am-Embryo_neg_3_compounds.tsv', quote=FALSE, sep='\t', col.names = NA)
```

```{r}
# create a sample info table

pos.filtered.sampleinfo <- data.frame(
  sample = colnames(merged.pos.filtered[,2:15]),
  Population = c("Molino","Molino","Molino","Molino","Molino","Pachón","Pachón","Pachón","Pachón","Pachón","Surface_fish","Surface_fish","Surface_fish","Surface_fish"),
  stringsAsFactors = FALSE  # This ensures that text columns are not converted to factors
)

pos.filtered.sampleinfo$Population <- factor(pos.filtered.sampleinfo$Population, levels = c("Surface_fish", "Pachón", "Molino"))

#write.table(pos.filtered.sampleinfo, file='input/sampleinfo/prep_normDE_outliersremoved_pos.sampleinfo.tsv', quote=FALSE, sep='\t', col.names = NA)

neg.filtered.sampleinfo <- data.frame(
  sample = colnames(merged.neg.filtered[,2:15]),
  Population = c("Molino","Molino","Molino","Molino","Molino","Pachón","Pachón","Pachón","Pachón","Pachón","Surface_fish","Surface_fish","Surface_fish","Surface_fish"),
  stringsAsFactors = FALSE  # This ensures that text columns are not converted to factors
)

neg.filtered.sampleinfo$Population <- factor(neg.filtered.sampleinfo$Population, levels = c("Surface_fish", "Pachón", "Molino"))

#write.table(neg.filtered.sampleinfo, file='input/sampleinfo/prep_normDE_outliersremoved_neg.sampleinfo.tsv', quote=FALSE, sep='\t', col.names = NA)
```

## 2.2 NormalyzerDE evaluation

### positive ion mode

Removed outliers group

```{r}
jobName <- "vignette_run_outliersremoved_pos"
dataFp <- "/Volumes/projects/fx2482/lipidomics-2cellstage/input/individual/prep_normDE_outliersremoved_Am-Embryo_pos_6_compounds.tsv"
designFp<- "/Volumes/projects/fx2482/lipidomics-2cellstage/input/sampleinfo/prep_normDE_outliersremoved_pos.sampleinfo.tsv"

experimentObj <- setupRawDataObject(dataFp, designFp, "default", TRUE, "sample", "Population")
normObj <- getVerifiedNormalyzerObject(jobName, experimentObj)

# 2. generate normalizations
normResults <- normMethods(normObj)

# 3. generate performance measures
normResultsWithEval <- analyzeNormalizations(normResults)

# 4. output matrices to file
jobDir <- setupJobDir("vignette_run_outliersremoved_pos", "/Volumes/projects/fx2482/lipidomics-2cellstage/output/")
#writeNormalizedDatasets(normResultsWithEval, jobDir)

# 5. generate evaluation plots
#generatePlots(normResultsWithEval, jobDir)

rm(jobDir,jobName, dataFp,designFp,experimentObj,normObj,normResults,normResultsWithEval)
```

### negative ion mode

Removed outliers group

```{r}
jobName <- "vignette_run_outliersremoved_neg"
dataFp <- "/Volumes/projects/fx2482/lipidomics-2cellstage/input/individual/prep_normDE_outliersremoved_Am-Embryo_neg_3_compounds.tsv"
designFp<- "/Volumes/projects/fx2482/lipidomics-2cellstage/input/sampleinfo/prep_normDE_outliersremoved_neg.sampleinfo.tsv"

experimentObj <- setupRawDataObject(dataFp, designFp, "default", TRUE, "sample", "Population")
normObj <- getVerifiedNormalyzerObject(jobName, experimentObj)

# 2. generate normalizations
normResults <- normMethods(normObj)

# 3. generate performance measures
normResultsWithEval <- analyzeNormalizations(normResults)

# 4. output matrices to file
jobDir <- setupJobDir("vignette_run_outliersremoved_neg", "/Volumes/projects/fx2482/lipidomics-2cellstage/output/")
#writeNormalizedDatasets(normResultsWithEval, jobDir)

# 5. generate evaluation plots
#generatePlots(normResultsWithEval, jobDir)

rm(jobDir,jobName, dataFp,designFp,experimentObj,normObj,normResults,normResultsWithEval)
```


# 3. Differential expression analysis

### positive ion mode - median normalization

```{r}
# 1. Setup folders and data matrices
jobDir <- "/Volumes/projects/fx2482/lipidomics-2cellstage/output/vignette_run_outliersremoved_pos/"
designFp<- "/Volumes/projects/fx2482/lipidomics-2cellstage/input/sampleinfo/prep_normDE_outliersremoved_pos.sampleinfo.tsv"

bestNormMatPath <- paste(jobDir, "median-normalized.txt", sep="/")
experimentObj <- setupRawContrastObject(bestNormMatPath, designFp, "sample")
nst <- NormalyzerStatistics(experimentObj, logTrans=FALSE)

# 2. Calculate statistics
# Now we are ready to perform the contrasts. Contrasts are provided as a vector in the format c("condA-condB", "condB-condC"), where condX is the group levels.

comparisons <- c("Molino-Surface_fish", "Pachón-Surface_fish")
nst <- calculateContrasts(nst, comparisons, condCol="Population", leastRepCount=2)

# 3. Generate final matrix and output
pos.Df <- generateAnnotatedMatrix(nst)
pos.jobDir <- setupJobDir("median-norm/DEG_analysis", "/Volumes/projects/fx2482/lipidomics-2cellstage/output/vignette_run_outliersremoved_pos/")

# load lipid names in the dataframe
test <- merged.pos.filtered
test$V1 <- as.integer(rownames(merged.pos.filtered))
pos.annotDf <- left_join(test[,c(1,16)], pos.Df,by = c("V1" = "NA") )
pos.annotDf <- pos.annotDf[,c(1,3:23)]

#write.csv(pos.annotDf, file=paste(pos.jobDir, "stat_table.csv", sep="/"))
#generateStatsReport(nst, "Vignette stats", pos.jobDir)

rm(jobDir,designFp,bestNormMatPath,experimentObj,nst,comparisons,test,pos.Df)
```

### negative ion mode - median normalization

```{r}
# 1. Setup folders and data matrices
jobDir <- "/Volumes/projects/fx2482/lipidomics-2cellstage/output/vignette_run_outliersremoved_neg/"
designFp<- "/Volumes/projects/fx2482/lipidomics-2cellstage/input/sampleinfo/prep_normDE_outliersremoved_neg.sampleinfo.tsv"

bestNormMatPath <- paste(jobDir, "median-normalized.txt", sep="/")
experimentObj <- setupRawContrastObject(bestNormMatPath, designFp, "sample")
nst <- NormalyzerStatistics(experimentObj, logTrans=FALSE)

# 2. Calculate statistics
# Now we are ready to perform the contrasts. Contrasts are provided as a vector in the format c("condA-condB", "condB-condC"), where condX is the group levels.

comparisons <- c("Molino-Surface_fish", "Pachón-Surface_fish")
nst <- calculateContrasts(nst, comparisons, condCol="Population", leastRepCount=2)

# 3. Generate final matrix and output
neg.Df <- generateAnnotatedMatrix(nst)
neg.jobDir <- setupJobDir("median-norm/DEG_analysis", "/Volumes/projects/fx2482/lipidomics-2cellstage/output/vignette_run_outliersremoved_neg/")

# load lipid names in the dataframe
test <- merged.neg.filtered
test$V1 <- as.integer(rownames(merged.neg.filtered))
neg.annotDf <- left_join(test[,c(1,16)], neg.Df,by = c("V1" = "NA") )
neg.annotDf <- neg.annotDf[,c(1,3:23)]

#write.csv(neg.annotDf, file=paste(neg.jobDir, "stat_table.csv", sep="/"))
#generateStatsReport(nst, "Vignette stats", neg.jobDir)

rm(jobDir,designFp,bestNormMatPath,experimentObj,nst,comparisons,test,neg.Df)
```


## 3.1 check if two collection methods are well separated

positive ion mode

```{r}

# Comparison as a whole

# comparison data creation

df <- merged.pos.filtered[,c(1,7:11)]


# Reshaping data to long format
long_df <- pivot_longer(df, cols = -Name, names_to = "Sample", values_to = "Value")
long_df <- long_df %>%
  mutate(Method = case_when(
    Sample %in% c("Group.Area..P1_pos", "Group.Area..P2_pos") ~ "NS", 
    Sample %in% c("Group.Area..P3_pos", "Group.Area..P5_pos", "Group.Area..P6_pos") ~ "IVF"
  ))

# Check normality with visual methods
ggplot(long_df, aes(sample = Value)) +
  facet_wrap(~ Method) +
  stat_qq() +
  stat_qq_line() +
  theme_minimal() +
  labs(title = "Q-Q Plots by Method") # They are not normally distributed

# Check homogeneity of variances (Levene's Test)
leveneTest(Value ~ Method, data = long_df) # homogeneous

# Perform Mann-Whitney U test
wilcox.test(Value ~ Method, data = long_df) # non-significant

rm(df,long_df)
```

```{r}
# Comparison individually

df <- merged.pos.filtered[,c(1,7:11)]

# Create an empty vector to store p-values
p_values <- numeric(nrow(df))

# Loop through each row/gene
for (i in 1:nrow(df)) {
  # Subset data for the current row/gene
  data_row <- df[i, ]
  
  # Subset data based on the method of collection
  data_method_ns <- data_row[colnames(data_row) %in% c("Group.Area..P1_pos", "Group.Area..P2_pos")]
  data_method_ivf <- data_row[colnames(data_row) %in% c("Group.Area..P3_pos", "Group.Area..P5_pos", "Group.Area..P6_pos")]
  
  # Perform t-test
  t_test_result <- t.test(data_method_ns, data_method_ivf)
  
  # Extract p-value and store it
  p_values[i] <- t_test_result$p.value
}

df$p.value <- p_values
df$p.adj <- p.adjust(p_values, method = "BH")

# print any significant expressed lipid names
print(df[df$p.adj < 0.05, ])

rm(df,p_values,i,data_row,data_method_ivf,data_method_ns,t_test_result)

```

negative ion mode

```{r}

# Comparison as a whole

# comparison data creation

df <- merged.neg.filtered[,c(1,7:11)]

# Reshaping data to long format
long_df <- pivot_longer(df, cols = -Name, names_to = "Sample", values_to = "Value")
long_df <- long_df %>%
  mutate(Method = case_when(
    Sample %in% c("Group.Area..P1_neg", "Group.Area..P2_neg") ~ "NS", 
    Sample %in% c("Group.Area..P3_neg", "Group.Area..P5_neg", "Group.Area..P6_neg") ~ "IVF"
  ))

# Check normality with visual methods
ggplot(long_df, aes(sample = Value)) +
  facet_wrap(~ Method) +
  stat_qq() +
  stat_qq_line() +
  theme_minimal() +
  labs(title = "Q-Q Plots by Method") # They are not normally distributed

by(long_df$Value, long_df$Method, shapiro.test) # not normally distributed

# Check homogeneity of variances (Levene's Test)
leveneTest(Value ~ Method, data = long_df) # homogeneous

# Perform Mann-Whitney U test
wilcox.test(Value ~ Method, data = long_df) # significant

rm(df,long_df)
```

```{r}
# Comparison individually

df <- merged.neg.filtered[,c(1,7:11)]

# Create an empty vector to store p-values
p_values <- numeric(nrow(df))

# Loop through each row/gene
for (i in 1:nrow(df)) {
  # Subset data for the current row/gene
  data_row <- df[i, ]
  
  # Subset data based on the method of collection
  data_method_ns <- data_row[colnames(data_row) %in% c("Group.Area..P1_neg", "Group.Area..P2_neg")]
  data_method_ivf <- data_row[colnames(data_row) %in% c("Group.Area..P3_neg", "Group.Area..P5_neg", "Group.Area..P6_neg")]
  
  # Perform t-test
  t_test_result <- t.test(data_method_ns, data_method_ivf)
  
  # Extract p-value and store it
  p_values[i] <- t_test_result$p.value
}

df$p.value <- p_values
df$p.adj <- p.adjust(df$p.value, method = "BH")

# print any significant expressed lipid names
print(df[df$p.adj < 0.05, ])

rm(i,df,p_values,data_row,data_method_ns,data_method_ivf,t_test_result)
```

No individual lipid species were found in negative ion mode. I will now assume it has not changed to the specific lipid level.

## 3.2 check for post-normalization matrix

positive ion mode

```{r}
cols<-rep(c("darkcyan","darkorange3"),each=3)
boxplot(pos.annotDf[,9:22],col=cols,main="after median normalization in positive ion mode")
colSums(pos.annotDf[,9:22])
```

```{r}
cor_plot <- pheatmap::pheatmap(cor(log2(pos.annotDf[,9:22])), main = "after median normalization in positive ion mode")
#ggsave("output/vignette_run_outliersremoved_pos/median-norm/plots/after-norm-correlation-plot.png", plot = cor_plot,width = 10, height = 10, units = "in")
```

negative ion mode

```{r}
boxplot(neg.annotDf[,9:22],col=cols,main="after median normalization in negative ion mode")
colSums(neg.annotDf[,9:22])
```

```{r}
cor_plot <- pheatmap::pheatmap(cor(log2(neg.annotDf[,9:22])), main = "after median normalization in negative ion mode")
#ggsave("output/vignette_run_outliersremoved_neg/median-norm/plots/after-norm-correlation-plot.png", plot = cor_plot,width = 10, height = 10, units = "in")

rm(cor_plot)
```

## 3.3 Annotation for lipid categories and InChIKey number

### positive ion mode

```{r}
# R-based tools for standardization of metabolite names (RefMet)
# https://github.com/metabolomicsworkbench/RefMet
# reference: https://rdcu.be/caRk5

# generate text file that the package requires
#write.table(pos.annotDf$Name, "output/vignette_run_outliersremoved_pos/lipids in positive ion mode.txt",quote = FALSE, row.names = FALSE, col.names = FALSE, sep = "\t") # mannually check whether lines are intact, add first line as a heading e.g. "NAME", "METABOLITE_NAME", "ID", etc


jobDir <- "/Volumes/projects/fx2482/lipidomics-2cellstage/output/vignette_run_outliersremoved_pos/lipids in positive ion mode.txt"
RefMet_mapped <- refmet_map(jobDir)
head(RefMet_mapped[,1:4])

RefMet_mapped$inchi_key <- numeric(nrow(RefMet_mapped))
# Use a data frame column containing metabolite names as input
for (i in 1:233){
  metadata <- refmet_metadata(RefMet_mapped[i, "Standardized.name"])
  RefMet_mapped$inchi_key[i] <- metadata[1,"inchi_key"]
} #line 234 has an error

for (i in 235:466){
  metadata <- refmet_metadata(RefMet_mapped[i, "Standardized.name"])
  RefMet_mapped$inchi_key[i] <- metadata[1,"inchi_key"]
} #line 467 has an error

metadata <- refmet_metadata(RefMet_mapped[468, "Standardized.name"])
  RefMet_mapped$inchi_key[468] <- metadata[1,"inchi_key"] #line 469 has an error

for (i in 470:673){
  metadata <- refmet_metadata(RefMet_mapped[i, "Standardized.name"])
  RefMet_mapped$inchi_key[i] <- metadata[1,"inchi_key"]
} #line 674 has an error
  
for (i in 675:nrow(RefMet_mapped)){
  metadata <- refmet_metadata(RefMet_mapped[i, "Standardized.name"])
  RefMet_mapped$inchi_key[i] <- metadata[1,"inchi_key"]
} #line 674 has an error  

#write.csv(RefMet_mapped, "output/vignette_run_outliersremoved_pos/reference_map.csv") 

# manually annotate most lipids through lipid maps
pos.RefMet_mapped <- read.csv("output/vignette_run_outliersremoved_pos/manual-reference_map.csv")

# combine this with DEG analysis
pos.RefMet_mapped <- left_join(pos.RefMet_mapped,pos.annotDf, by = c("Input.name" = "Name"))
rownames(pos.RefMet_mapped[is.na(pos.RefMet_mapped$`Molino-Surface_fish_PValue`),])

#There are a few NAs need to be corrected
pos.RefMet_mapped[3,1] <- pos.annotDf[3,1]
pos.RefMet_mapped[3, 9:29] <- pos.annotDf[3, 2:22]

pos.RefMet_mapped[27,1] <- pos.annotDf[27,1]
pos.RefMet_mapped[27, 9:29] <- pos.annotDf[27, 2:22]

pos.RefMet_mapped[184,1] <- pos.annotDf[184,1]
pos.RefMet_mapped[184, 9:29] <- pos.annotDf[184, 2:22]

rownames(pos.RefMet_mapped[is.na(pos.RefMet_mapped$`Molino-Surface_fish_PValue`),])# no NAs

# delete mass and formula column
pos.RefMet_mapped <- pos.RefMet_mapped[, c(1:2, 5:29)]
#write.csv(pos.RefMet_mapped, "output/vignette_run_outliersremoved_pos/median-norm/DEG_analysis/all lipids in positive ion mode.csv")

rm(pos.jobDir)
```


### negative ion mode

```{r}
# R-based tools for standardization of metabolite names (RefMet)
# https://github.com/metabolomicsworkbench/RefMet
# reference: https://rdcu.be/caRk5

# generate text file that the package requires
#write.table(neg.annotDf$Name, "output/vignette_run_outliersremoved_neg/lipids in negative ion mode.txt",quote = FALSE, row.names = FALSE, col.names = FALSE, sep = "\t") # mannually check whether lines are intact, add first line as a heading e.g. "NAME", "METABOLITE_NAME", "ID", etc

jobDir <- "/Volumes/projects/fx2482/lipidomics-2cellstage/output/vignette_run_outliersremoved_neg/lipids in negative ion mode.txt"
RefMet_mapped <- refmet_map(jobDir)
head(RefMet_mapped[,1:4])

RefMet_mapped$inchi_key <- numeric(nrow(RefMet_mapped))
# Use a data frame column containing metabolite names as input
for (i in 1:28){
  metadata <- refmet_metadata(RefMet_mapped[i, "Standardized.name"])
  RefMet_mapped$inchi_key[i] <- metadata[1,"inchi_key"]
} #line 29 has an error

for (i in 30:131){
  metadata <- refmet_metadata(RefMet_mapped[i, "Standardized.name"])
  RefMet_mapped$inchi_key[i] <- metadata[1,"inchi_key"]
} #line 132 and 133 has an error

for (i in 134:142){
  metadata <- refmet_metadata(RefMet_mapped[i, "Standardized.name"])
  RefMet_mapped$inchi_key[i] <- metadata[1,"inchi_key"]
} #line 143 has an error

for (i in 144:236){
  metadata <- refmet_metadata(RefMet_mapped[i, "Standardized.name"])
  RefMet_mapped$inchi_key[i] <- metadata[1,"inchi_key"]
} #line 237 has an error

for (i in 238:670){
  metadata <- refmet_metadata(RefMet_mapped[i, "Standardized.name"])
  RefMet_mapped$inchi_key[i] <- metadata[1,"inchi_key"]
} #line 671 has an error

for (i in 672:690){
  metadata <- refmet_metadata(RefMet_mapped[i, "Standardized.name"])
  RefMet_mapped$inchi_key[i] <- metadata[1,"inchi_key"]
} #line 691 has an error

  
for (i in 692:nrow(RefMet_mapped)){
  metadata <- refmet_metadata(RefMet_mapped[i, "Standardized.name"])
  RefMet_mapped$inchi_key[i] <- metadata[1,"inchi_key"]
} #line 691 has an error  

#write.csv(RefMet_mapped, "output/vignette_run_outliersremoved_neg/reference_map.csv") # need a lot of manual annotation

# manually annotate most lipids through lipid maps
neg.RefMet_mapped <- read.csv("output/vignette_run_outliersremoved_neg/manual-reference_map.csv")

# combine this with DEG analysis
neg.RefMet_mapped <- left_join(neg.RefMet_mapped,neg.annotDf, by = c("Input.name" = "Name"))
rownames(pos.RefMet_mapped[is.na(pos.RefMet_mapped$`Molino-Surface_fish_PValue`),])

#There are a few NAs need to be corrected
neg.RefMet_mapped[2,1] <- neg.annotDf[2,1]
neg.RefMet_mapped[2, 9:29] <- neg.annotDf[2, 2:22]

neg.RefMet_mapped[3,1] <- neg.annotDf[3,1]
neg.RefMet_mapped[3, 9:29] <- neg.annotDf[3, 2:22]

neg.RefMet_mapped[6,1] <- neg.annotDf[6,1]
neg.RefMet_mapped[6, 9:29] <- neg.annotDf[6, 2:22]

neg.RefMet_mapped[718,1] <- neg.annotDf[718,1]
neg.RefMet_mapped[718, 9:29] <- neg.annotDf[718, 2:22]

neg.RefMet_mapped[730,1] <- neg.annotDf[730,1]
neg.RefMet_mapped[730, 9:29] <- neg.annotDf[730, 2:22]

rownames(neg.RefMet_mapped[is.na(neg.RefMet_mapped$`Molino-Surface_fish_PValue`),])# no NAs

# delete mass and formula column
neg.RefMet_mapped <- neg.RefMet_mapped[, c(1:2, 5:29)]
#write.csv(neg.RefMet_mapped, "output/vignette_run_outliersremoved_neg/median-norm/DEG_analysis/all lipids in negative ion mode.csv")

rm(neg.jobDir)

```



## 3.4 results from DEG analysis

### positive ion mode
```{r}
jobDir <- "/Volumes/projects/fx2482/lipidomics-2cellstage/output/vignette_run_outliersremoved_pos/median-norm/DEG_analysis/"

pos.up_cav_mapped <- pos.RefMet_mapped[pos.RefMet_mapped$`Molino-Surface_fish_AdjPVal` < 0.05 & pos.RefMet_mapped$`Pachón-Surface_fish_AdjPVal` < 0.05 & pos.RefMet_mapped$`Molino-Surface_fish_log2FoldChange` > 0.5 & pos.RefMet_mapped$`Pachón-Surface_fish_log2FoldChange` >0.5, ]

pos.down_cav_mapped <- pos.RefMet_mapped[pos.RefMet_mapped$`Molino-Surface_fish_AdjPVal` < 0.05 & pos.RefMet_mapped$`Pachón-Surface_fish_AdjPVal` < 0.05 & pos.RefMet_mapped$`Molino-Surface_fish_log2FoldChange` < -0.5 & pos.RefMet_mapped$`Pachón-Surface_fish_log2FoldChange` < -0.5, ]

#write.csv(pos.up_cav_mapped, file=paste(jobDir, "mapped lipids upregulated in cavefish in positive ion mode.csv", sep="/"))
#write.csv(pos.down_cav_mapped, file=paste(jobDir, "mapped lipids downregulated in cavefish in positive ion mode.csv", sep="/"))

rm(jobDir)
```

### negative ion mode
```{r}
jobDir <- "/Volumes/projects/fx2482/lipidomics-2cellstage/output/vignette_run_outliersremoved_neg/median-norm/DEG_analysis/"

neg.up_cav_mapped <- neg.RefMet_mapped[neg.RefMet_mapped$`Molino-Surface_fish_AdjPVal` < 0.05 & neg.RefMet_mapped$`Pachón-Surface_fish_AdjPVal` < 0.05 & neg.RefMet_mapped$`Molino-Surface_fish_log2FoldChange` > 0.5 & neg.RefMet_mapped$`Pachón-Surface_fish_log2FoldChange` >0.5, ]

neg.down_cav_mapped <- neg.RefMet_mapped[neg.RefMet_mapped$`Molino-Surface_fish_AdjPVal` < 0.05 & neg.RefMet_mapped$`Pachón-Surface_fish_AdjPVal` < 0.05 & neg.RefMet_mapped$`Molino-Surface_fish_log2FoldChange` < -0.5 & neg.RefMet_mapped$`Pachón-Surface_fish_log2FoldChange` < -0.5, ]

#write.csv(neg.up_cav_mapped, file=paste(jobDir, "mapped lipids upregulated in cavefish in negitive ion mode.csv", sep="/"))
#write.csv(neg.down_cav_mapped, file=paste(jobDir, "mapped lipids downregulated in cavefish in negitive ion mode.csv", sep="/"))

rm(jobDir)
```



```{r}
rm(merged.neg,merged.neg.filtered,merged.pos,merged.pos.filtered,neg.annotDf,neg.down_cav_mapped,neg.filtered.sampleinfo,neg.RefMet_mapped,neg.up_cav_mapped,pos.annotDf,pos.down_cav_mapped,pos.filtered.sampleinfo,pos.RefMet_mapped,pos.up_cav_mapped)
```


# Session Info
```{r}
sessionInfo()
```
