---
title: "proteomics-analysis"
author: "Fanning Xia"
output: html_document
editor_options: 
  chunk_output_type: console
---

# 0. Preparation work
## Load, filter the data and create sample info

```{r}
# load packages
library(NormalyzerDE)
library(preprocessCore)
library(ggplot2)
library(dplyr)
library(UniprotR)
library(rentrez)
library(reshape2)
library(dendextend)
library(biomaRt)
library(ggfortify)
library(cluster)
library(tidyr)
library(ggsci)
library(scales)
library(car)
library(FSA)
library(svglite)

pvg3 <- c("#979FA2","#EE9B00","#93D2DF")
```

```{r}
# set working directory
setwd(".../proteomics-2cellstage/")

# load data for heatmap
counts <- read.csv("input/raw_norm_Am_cavefish-embryo_WCE_PROT941-TMT9_hph_neo_125min_2.csv")
counts <- counts[, c(4:5,8,30:66)]

# remove Contaminant
valid_counts <- counts[!grepl("^Contaminant_", counts$Accession),]

# remove other parameters, just keep raw counts
rownames(valid_counts) <- valid_counts[,1]

ms_normcounts <- valid_counts[,23:31] # normalized counts from Mass Spec core
valid_counts <- valid_counts[,14:22] # raw counts

# check if there's any NA not in the entire row
rows_with_partial_NAs <- valid_counts[rowSums(is.na(valid_counts)) > 0 & rowSums(!is.na(valid_counts)) > 0, ]
rows_with_partial_NAs #results show no partial NAs, we can remove this parameter
rm(rows_with_partial_NAs)

# remove all lines with NA 
valid_counts <- na.omit(valid_counts)
ms_normcounts <- na.omit(ms_normcounts)

# find unique rows, need to remove the rows with exact same value in every columns (remove repetitive rows)
# Initally I want to use PEP score to keep unique rows, but PEP score is not necessarily correlated with unique names
# unique_counts <- counts[!duplicated(counts$Sum.PEP.Score), ]

unique_counts <- unique(valid_counts)
ms_normcounts <- unique(ms_normcounts)

#unique_counts[unique_counts ==  0] <- 1 # was trying to change it to 0.01, but it will have an error during log normalization. I will check eventually if there is one gene differentially expressed with this artifical 1

#write.table(unique_counts, file='input/prep_normDE_counts.tsv', quote=FALSE, sep='\t', col.names = NA)
#write.table(rownames(unique_counts)[grep("^XP|^NP", rownames(unique_counts))], file='input/prep_normDE_XP_ID.tsv', quote=FALSE, sep='\t', col.names = NA)
#write.table(rownames(unique_counts)[grep("^KAG", rownames(unique_counts))], file='input/prep_normDE_KAG_ID.tsv', quote=FALSE, sep='\t', col.names = NA)


# create a sample info table
sampleinfo <- data.frame(
  sample = colnames(unique_counts),
  group = c("Molino","Molino","Molino", "Pachón","Pachón","Pachón", "Surface_fish","Surface_fish","Surface_fish"),
  stringsAsFactors = FALSE  # This ensures that text columns are not converted to factors
)

#write.table(sampleinfo, file='input/prep_normDE_sampleinfo.tsv', quote=FALSE, sep='\t', col.names = NA)

```


# 1. Normalization methods

Used NormalyzerDE

Paper: NormalyzerDE: Online Tool for Improved Normalization of Omics Expression Data and High-Sensitivity Differential Expression Analysis
https://pubs.acs.org/doi/10.1021/acs.jproteome.8b00523

Package: Evaluation and statistics of expression data using NormalyzerDE
https://www.bioconductor.org/packages/release/bioc/vignettes/NormalyzerDE/inst/doc/vignette.html#stepwise-processing-normalization-part

## 1.1 Running NormalyzerDE evaluation

```{r}
# 1. load data
jobName <- "vignette_run"
designFp <- "/Volumes/projects/fx2482/proteomics-2cellstage/input/prep_normDE_sampleinfo.tsv"
dataFp <- "/Volumes/projects/fx2482/proteomics-2cellstage/input/prep_normDE_counts.tsv"

experimentObj <- setupRawDataObject(dataFp, designFp, "default", TRUE, "sample", "group")
normObj <- getVerifiedNormalyzerObject(jobName, experimentObj)

# 2. generate normalizations
normResults <- normMethods(normObj)

# 3. generate performance measures
normResultsWithEval <- analyzeNormalizations(normResults)

# 4. output matrices to file
jobDir <- setupJobDir("vignette_run", "/Volumes/projects/fx2482/proteomics-2cellstage/output/")
#writeNormalizedDatasets(normResultsWithEval, jobDir)

# 5. generate evaluation plots
#generatePlots(normResultsWithEval, jobDir)

rm(jobName, dataFp, experimentObj,normObj,normResults,normResultsWithEval)

```

## 1.2 Median normalization - Differential expression analysis

```{r}
# 1. Setup folders and data matrices

bestNormMatPath <- paste(jobDir, "median-normalized.txt", sep="/")
experimentObj <- setupRawContrastObject(bestNormMatPath, designFp, "sample")
nst <- NormalyzerStatistics(experimentObj, logTrans=FALSE)

# 2. Calculate statistics
# Now we are ready to perform the contrasts. Contrasts are provided as a vector in the format c("condA-condB", "condB-condC"), where condX is the group levels.

comparisons <- c("Molino-Surface_fish", "Pachón-Surface_fish")
nst <- calculateContrasts(nst, comparisons, condCol="group", leastRepCount=2)

# 3. Generate final matrix and output
annotDf <- generateAnnotatedMatrix(nst)
jobDir <- setupJobDir("DEG-analysis", ".../proteomics-2cellstage/output/vignette_run/median-norm/")
#write.csv(annotDf, file=paste(jobDir, "stat_table.csv", sep="/"))
#generateStatsReport(nst, "Vignette stats", jobDir)

rm(bestNormMatPath,experimentObj,nst,comparisons)
```

## 1.3 Annotation: transfer protein ID to entrez ID (NCBI gene ID)
version (not entirely sure, mostly should be AstMex_SF_3)
https://www.ncbi.nlm.nih.gov/datasets/genome/?taxon=7994


### 1.3.1 For XP protein ids - rentrez

```{r}
# load rentrez package to interact with the NCBI E-utilities
library(rentrez)

# Define a vector of protein IDs
protein_ids <- rownames(unique_counts)[grep("^XP|^NP", rownames(unique_counts))]

# Define the database to query (e.g., 'gene' for Entrez Gene)
database <- "gene"

# Function to convert protein IDs to Entrez Gene IDs
convert_protein_ids <- function(protein_ids, database = "gene") {
  # Initialize a list to store results
  gene_ids <- vector("list", length = length(protein_ids))
  
  # Loop over each protein ID
  for (i in seq_along(protein_ids)) {
    # Construct the query
    query <- paste(protein_ids[i], "[Accession]", sep = " ")
    
    # Perform the search
    search_results <- entrez_search(db = database, term = query)
    
    # Extract the Entrez Gene ID from the search results
    if (search_results$count > 0) {
      gene_ids[[i]] <- search_results$ids[1]
    } else {
      gene_ids[[i]] <- NA
    }
  }
  
  # Return the list of Entrez Gene IDs
  return(gene_ids)
}

# Convert protein IDs to Entrez Gene IDs
entrez_gene_ids <- convert_protein_ids(protein_ids, database) 

# Print the results
for (i in seq_along(protein_ids)) {
  cat("Protein ID:", protein_ids[i], "corresponds to Entrez Gene ID:", entrez_gene_ids[[i]], "\n")
}

# get the unlist result
annotDf_gene <- data.frame(
  protein_id = protein_ids,
  gene_id = unlist(entrez_gene_ids),
  stringsAsFactors = FALSE  # This ensures that text columns are not converted to factors
)

#write.csv(annotDf_gene,"output/vignette_run/annotation/XPandNP_protein_to_gene_annotation.csv", sep="/"))

rm(protein_ids,database)

```

### 1.3.2 For KAG protein ids

```{r}

# first get old mapping from MS core / ThermoFisher

counts <- read.csv("input/raw_norm_Am_cavefish-embryo_WCE_PROT941-TMT9_hph_neo_125min_2.csv")
rownames(counts) <- counts$Accession

protein_ids_KAG <- rownames(unique_counts)[grep("^KAG", rownames(unique_counts))]
protein_KAG <- counts[protein_ids_KAG,]

test <- protein_KAG[,c(4,22,24,25)]
test$Gene.Symbol <- tolower(test$Gene.Symbol) #get them into lower case


```

#### 1.3.2.1 Use biomaRt to find gene ids

```{r}

ensembl <- useEnsembl(biomart = "ensembl",
                       dataset = "amexicanus_gene_ensembl",
                       mirror = "useast")

filters <- listFilters(ensembl)
grep("XP_", filters$description, value = TRUE, ignore.case = TRUE)
grep("KAG", filters$description, value = TRUE, ignore.case = TRUE)
grep("ENS", filters$description, value = TRUE, ignore.case = TRUE)

test.1 <- getBM(attributes = c("entrezgene_id","entrezgene_accession"),
                          filters = "entrezgene_accession",
                          values = test$Gene.Symbol,
                          mart = ensembl)
#test.2 <- left_join(test, test.1, by = c("Gene.Symbol" = "entrezgene_accession"))

# then ony keep one lane once we found each conditions
# find common annotated rows between pd annotation and our annotation
#test.3 <- test.2[test.2$Entrez.Gene.ID == test.2$entrezgene_id,]
#test.3 <- na.omit(test.3)
#test.3 <- test.3[,c(1:4)]

# find annotated rows only in new mapping
#test.4 <- test.2[test.2$Entrez.Gene.ID == "" & !is.na(test.2$entrezgene_id),]
#test.4$Entrez.Gene.ID <- test.4$entrezgene_id
#test.4 <- test.4[,c(1:4)]

# find annotated rows only in original mapping
#test.5 <- test.2[!test.2$Entrez.Gene.ID == "" & is.na(test.2$entrezgene_id),]
#test.5 <- test.5[,c(1:4)]

# find annotated rows with different values
#test.6 <- test.2[!test.2$Entrez.Gene.ID == "" & !is.na(test.2$entrezgene_id) & test.2$Entrez.Gene.ID != test.2$entrezgene_id,]
#test.6 <- test.6[,c(1:4)] #by manual check, originally mapped ones were correct (mostly)

# all unannotated ones
#test.7 <- test.2[test.2$Entrez.Gene.ID == "" & is.na(test.2$entrezgene_id),]
#test.7 <- test.7[,c(1:4)]

#annoDF_protein_KAG <- rbind(test.3,test.4,test.5,test.6)
#rm(test,test.1,test.2,test.3,test.4,test.5,test.6)
```

#### 1.3.2.2 Use rentrez to find gene ids

```{r}
# Define a vector of leftover genes unannotated in test.7
#gene_names <- test.7$Gene.Symbol

# Define the database to query (e.g., 'gene' for Entrez Gene)
#database <- "gene"

# Function to convert pd annotated gene names IDs to Entrez Gene IDs
convert_gene_names <- function(gene_names, database = "gene") {
  # Initialize a list to store results
  gene_ids <- vector("list", length = length(gene_names))
  
  # Loop over each gene_names
  for (i in seq_along(gene_names)) {
    # Construct the query
    query <- paste(gene_names[i], "[Gene Name] AND", 7994, "[Taxonomy ID]", sep = " ")
    
    # Perform the search
    search_results <- entrez_search(db = database, term = query)
    
    # Extract the Entrez Gene ID from the search results
    if (search_results$count > 0) {
      gene_ids[[i]] <- search_results$ids[1]
    } else {
      gene_ids[[i]] <- NA
    }
  }
  
  # Return the list of Entrez Gene IDs
  return(gene_ids)
}

# Convert protein IDs to Entrez Gene IDs
#geneid_based_on_gene_names <- convert_gene_names(gene_names, database)

#test.7$Entrez.Gene.ID <- unlist(geneid_based_on_gene_names)

# combine it 
#annoDF_protein_KAG <- rbind(annoDF_protein_KAG, test.7[complete.cases(test.7) & test.7$Gene.Symbol != "",])

#rm(database)
```

#### 1.3.2.3 Manual annotation

```{r}
# find final rows without any annotations
#test.8 <- test.7[!complete.cases(test.7) | test.7$Gene.Symbol == "",]

#mannual annotation
#test.8$Entrez.Gene.ID <- c("103033944", "103045882", "103046345","103034887","111190432",
#                           "103029774", "","103047682","103022465","103046426",
#                           "103039390", "103028412", "107197318", "103025840","103038803",
#                           "111192975", "103034887", "103039618","103045988","103046720",
#                           "111191983", "103037668", "103042293", "103035393", "103023629",
#                           "103022532", "103027693", "103034296", "103036175", "103033541",
#                           "103040660", "103025109", "103034086", "103027409", "103022177",
#                           "103047533", "103037884", "103041583", "103024822", "103034233",
#                           "103023892", "107197708", "","103028044", "125806057",
#                           "111188250", "103046166", "103024139", "103036797", "103022403",
#                           "103041044", "103022710", "111193015", "103026176", "103026776",
#                           "111196314", "111197585", "103028468", "111191983", "103024037",
#                           "103026173", "103045882", "103046193", "103029047", "103028685",
#                           "111195933"
#                           )
#annoDF_protein_KAG <- rbind(annoDF_protein_KAG, test.8)
#write.csv("output/vignette_run/annotation/KAG_protein_to_gene_annotation.csv", sep="/"))

#rm(test.7,test.8)
```

### 1.3.3 Get the full annotation for all unique protein findings

```{r}
# make all annotations regardless of XP_ or KAG the same
# load annotations

ann_XP <- read.csv("output/vignette_run/annotation/XPandNP_protein_to_gene_annotation.csv")
ann_KAG <- read.csv("output/vignette_run/annotation/KAG_protein_to_gene_annotation.csv")

ann_XP <- ann_XP[,2:3]
ann_KAG <- ann_KAG[,2:3]
colnames(ann_KAG) <- colnames(ann_XP)

anno_all <- rbind(ann_KAG,ann_XP)

#write.csv(anno_all, "output/vignette_run/annotation/unique_protein_to_gene_annotation.csv")

library(dplyr)
annotDf_df <- left_join(anno_all, annotDf, c("protein_id"="V1"))
annotDf_df$gene_id <- as.integer(annotDf_df$gene_id)

#write.csv(annotDf_df, "output/vignette_run/median-norm/DEG-analysis/All_protein_to_gene_annotation.csv")

rm(ann_XP,ann_KAG,anno_all)

```

# 2. Check if two collection methods are well separated

All surface fish samples were collected through "natural spawning (NS)", and all Molino cavefish samples were collected through night IVF. In Pachón samples, 128N was collected by NS and 128C and 129N were by IVF. This is to test if the collection methods would affect the comparisons especially for Molino and surface fish.

```{r}
# Comparison as a whole

# comparison data creation

df <- annotDf_df[,c(2,13:15)]

# Reshaping data to long format
long_df <- pivot_longer(df, cols = -gene_id, names_to = "Sample", values_to = "Value")
long_df <- long_df %>%
  mutate(Method = case_when(
    Sample %in% "Abundance..F8..128N..Sample..Pachon" ~ "NS", 
    Sample %in% c("Abundance..F8..128C..Sample..Pachon", "Abundance..F8..129N..Sample..Pachon") ~ "IVF"
  ))

# Check normality for each group
ggplot(long_df, aes(sample = Value)) +
  facet_wrap(~ Method) +
  stat_qq() +
  stat_qq_line() +
  theme_minimal() +
  labs(title = "Q-Q Plots by Method") # They are not normally distributed

by(long_df$Value, long_df$Method, shapiro.test) # not normally distributed

# Check homogeneity of variances (Levene's Test)
leveneTest(Value ~ Method, data = long_df) # not homogeneous

# Perform Mann-Whitney U test
wilcox.test(Value ~ Method, data = long_df) # non-significant

rm(df,long_df)
```

```{r}
# Comparison individually

df <- annotDf_df[,c(2,13:15)]
df <- na.omit(df)

# Create an empty vector to store p-values
p_values <- numeric(nrow(df))

# Loop through each row/gene
for (i in 1:nrow(df)) {
  # Subset data for the current row/gene
  data_row <- df[i, ]
  
  # Subset data based on the method of collection
  data_method_ns <- data_row[colnames(data_row) == "Abundance..F8..128N..Sample..Pachon"]
  data_method_ivf <- data_row[colnames(data_row) %in% c("Abundance..F8..128C..Sample..Pachon", "Abundance..F8..129N..Sample..Pachon")]
  
  # Perform t-test
  t_test_result <- t.test(data_method_ivf, mu = data_method_ns$Abundance..F8..128N..Sample..Pachon)
  
  # Extract p-value and store it
  p_values[i] <- t_test_result$p.value
}

df$p.value <- p_values
df$p.adj <- p.adjust(df$p.value, method = "BH")

# print any significant expressed gene names
print(df[df$p.adj < 0.05, ])

rm(i,df,p_values,data_row,data_method_ns,data_method_ivf,t_test_result)
```

Conclusion for this: generally, the collection method might not impact the results. Only 2 genes were found significantly affected by the methods. I will need to keep these two in mind if anything pop up in the final result. They are: "eif3hb" and "gemin7"

# 3. Get results from DEG analysis

## Some parameters

```{r}
# How many unique gene ID?
length(unique(annotDf_df$gene_id))
```

```{r}
# How many percentage of vtgs? (an estimate)

# Find all vtgs (used ms annotations, I need to talk to Di about all annotations of shinyapps)

df <- annotDf_df
rownames(df) <- df$protein_id
df <- df[,10:18]
df = 2^df
df[is.na(df)] <- 0

sum(df[rownames(df) %in% c("XP_022535137.2","XP_049320372.1","XP_049321622.1","XP_022525474.2","XP_049321621.1","KAG9266789.1","XP_049321630.1","KAG9266788.1","XP_007233218.3","KAG9266776.1","XP_022532042.2"), ])/sum(df)
```

## DEG analysis

```{r}
up_cav <- annotDf_df[annotDf_df$`Molino-Surface_fish_AdjPVal` < 0.05 & annotDf_df$`Pachón-Surface_fish_AdjPVal` < 0.05 & annotDf_df$`Molino-Surface_fish_log2FoldChange` > 0.5 & annotDf_df$`Pachón-Surface_fish_log2FoldChange` >0.5, ]

up_pa <- annotDf_df[annotDf_df$`Pachón-Surface_fish_AdjPVal` < 0.05 & annotDf_df$`Pachón-Surface_fish_log2FoldChange` >0.5, ]

up_mo <- annotDf_df[annotDf_df$`Molino-Surface_fish_AdjPVal` < 0.05 & annotDf_df$`Molino-Surface_fish_log2FoldChange` > 0.5, ]

down_cav <- annotDf_df[annotDf_df$`Molino-Surface_fish_AdjPVal` < 0.05 & annotDf_df$`Pachón-Surface_fish_AdjPVal` < 0.05 & annotDf_df$`Molino-Surface_fish_log2FoldChange` < -0.5 & annotDf_df$`Pachón-Surface_fish_log2FoldChange` < -0.5, ]

down_pa <- annotDf_df[annotDf_df$`Pachón-Surface_fish_AdjPVal` < 0.05 & annotDf_df$`Pachón-Surface_fish_log2FoldChange` < -0.5, ]

down_mo <- annotDf_df[annotDf_df$`Molino-Surface_fish_AdjPVal` < 0.05 & annotDf_df$`Molino-Surface_fish_log2FoldChange` < -0.5, ]

#write.csv(up_cav, file=paste(jobDir, "annotated with NCBI gene id_upregulated in cavefish.csv", sep="/"))
#write.csv(down_cav, file=paste(jobDir, "annotated with NCBI gene id_downregulated in cavefish.csv", sep="/"))
#write.csv(up_pa, file=paste(jobDir, "annotated with NCBI gene id_upregulated in Pachón.csv", sep="/"))
#write.csv(up_mo, file=paste(jobDir, "annotated with NCBI gene id_upregulated in Molino.csv", sep="/"))
#write.csv(down_pa, file=paste(jobDir, "annotated with NCBI gene id_downregulated in Pachón.csv", sep="/"))
#write.csv(down_mo, file=paste(jobDir, "annotated with NCBI gene id_downregulated in Molino.csv", sep="/"))


rm(up_cav,down_cav,up_pa,up_mo,down_pa,down_mo)
```

## NEED TO DO MOREannotate gene name and use this for zebrafish GO term

```{r}

library(dplyr)

#annotDf_pro <- down_cav
#annotDf_pro <- na.omit(annotDf_pro)

#annotdf_protein_to_gene <- getBM(attributes = c("entrezgene_id","entrezgene_accession"),
#                          filters = "entrezgene_id",
#                          values = annotDf_pro$gene_id,
#                          mart = ensembl)

#annotDf_df <- left_join(annotdf_protein_to_gene, annotDf_test, by = c("refseq_peptide_predicted"="V1"))


#write.csv(annotdf_protein_to_gene, file=paste(jobDir, "downcav_gene_ann.csv", sep="/"))
```



# 4. Plots

## 4.1 Check for before and after normalization matrix boxplot

### 4.1.1 Boxplot

```{r}
cols<-rep(c("darkcyan","darkorange3"),each=3)
boxplot(log2(unique_counts),col=cols,main="before norm")
boxplot(log2(ms_normcounts),col=cols,main="after norm from Mass spec core facility") # PD used normalization to the whole intensity, similar to global intensity normalization
boxplot(annotDf_df[,10:18],col=cols,main="after median normalization")
```

### 4.1.2 Correlation plot

```{r}
# before normalization
temp <- na.omit(unique_counts)
temp[temp ==  0] <- 0.01
pheatmap::pheatmap(cor(log2(temp)), main = "before norm")
cor_plot <- pheatmap::pheatmap(cor(log2(temp)), main = "before norm")

#ggsave("output/vignette_run/median-norm/before-norm-correlation-plot.png", plot = cor_plot,width = 10, height = 10, units = "in")

# after norm from Mass spec core facility
temp <- na.omit(ms_normcounts)
temp[temp ==  0] <- 0.01
pheatmap::pheatmap(cor(log2(temp)), main = "after norm from Mass spec core facility", method = "spearman")

# after median normalization
temp <- na.omit(annotDf[,9:17])
pheatmap::pheatmap(cor(log2(temp)), main = "after median normalization")
cor_plot <- pheatmap::pheatmap(cor(log2(temp)), main = "after median normalization")

#ggsave("output/vignette_run/median-norm/median-norm-correlation-plot.png", plot = cor_plot,width = 10, height = 10, units = "in")
rm(temp, cor_plot)
```

### Figure 2B PCA plot

```{r}
#from https://cran.r-project.org/web/packages/ggfortify/vignettes/plot_pca.html
#from https://bioinformatics-core-shared-training.github.io/Bulk_RNAseq_Course_2021/Markdowns/07_Data_Exploration.html#principal-component-analysis

#To run the PCA we should first normalise our data for library size and transform to a log scale. 

# read the norm data
normCounts <- annotDf[,9:17]

# remove NA lines
filtered_normCounts <- normCounts[complete.cases(normCounts), ]

log_data <- log2(filtered_normCounts + 1)
pca_res <- prcomp(t(log_data), scale. = TRUE)
pca_plot <- autoplot(pca_res, label = TRUE)
#ggsave("output/vignette_run/median-norm/Proteomics - PCA plot (median-norm) with specific sample name.png", plot = pca_plot, width = 10, height = 8, units = "in")

rm(pca_plot)

# read sampleinfo
sampleinfo <- data.frame(
  sample = colnames(unique_counts),
  Population = c("Molino","Molino","Molino", "Pachón","Pachón","Pachón", "Surface fish","Surface fish","Surface fish"),
  stringsAsFactors = FALSE  # This ensures that text columns are not converted to factors
)

sampleinfo$Population <- factor(sampleinfo$Population, levels = c("Surface fish", "Pachón", "Molino"))

pca_plot <- autoplot(pca_res, data = sampleinfo,  colour='Population', size=1.8
                     #, frame = TRUE, frame.type = 'norm'
                     )

pca_plot <- pca_plot + 
  scale_color_manual(values = pvg3)+
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Add significance threshold
  geom_vline(xintercept = 0, linetype = "dashed") + # Add foldchange threshold
  theme_light()+
  labs(title = "Proteomics - PCA plot") +
  theme(text = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 0.5))  # Center the title

pca_plot
#Too few points to calculate an ellipse 

#ggsave("output/vignette_run/median-norm/paper/Figure 2B Proteomics - PCA plot (median-norm).png", plot = pca_plot, width = 6, height = 4, units = "in")
#ggsave("output/vignette_run/median-norm/paper/Figure 2B Proteomics - PCA plot (median-norm).svg", plot = pca_plot, width = 6, height = 4, units = "in")

rm(normCounts,filtered_normCounts,log_data,pca_res,pca_plot,sampleinfo)
```


## SuppFig 2A Volcano plot

```{r}
vol_plot_pa_sf <-ggplot(annotDf_df, aes(x = `Pachón-Surface_fish_log2FoldChange`, y = -log10(`Pachón-Surface_fish_AdjPVal`))) +
  geom_point(color = ifelse(
    annotDf_df$`Pachón-Surface_fish_AdjPVal` < 0.05 &
      annotDf_df$`Molino-Surface_fish_AdjPVal` < 0.05 & 
      (annotDf_df$`Pachón-Surface_fish_log2FoldChange` > 0.5 &
      annotDf_df$`Molino-Surface_fish_log2FoldChange` > 0.5) |
      annotDf_df$`Pachón-Surface_fish_AdjPVal` < 0.05 &
      annotDf_df$`Molino-Surface_fish_AdjPVal` < 0.05 &
      (annotDf_df$`Pachón-Surface_fish_log2FoldChange` < -0.5 &
      annotDf_df$`Molino-Surface_fish_log2FoldChange` < -0.5), "red",
    "grey"), alpha = 0.6) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "black") +  # Add significance threshold
  geom_vline(xintercept = c(-0.5, 0.5), linetype = "dashed") + # Add foldchange threshold
  theme_light()+
  labs(x = "Log2 Fold Change", y = "-log10(P adjusted)", title = "Pachón vs Surface fish") +
  theme(plot.title = element_text(hjust = 0.5)) 

vol_plot_pa_sf

vol_plot_mo_sf <- ggplot(annotDf_df, aes(x = `Molino-Surface_fish_log2FoldChange`, y = -log10(`Molino-Surface_fish_AdjPVal`))) +
  geom_point(color = ifelse(
    annotDf_df$`Pachón-Surface_fish_AdjPVal` < 0.05 &
      annotDf_df$`Molino-Surface_fish_AdjPVal` < 0.05 & 
      (annotDf_df$`Pachón-Surface_fish_log2FoldChange` > 0.5 &
      annotDf_df$`Molino-Surface_fish_log2FoldChange` > 0.5) |
      annotDf_df$`Pachón-Surface_fish_AdjPVal` < 0.05 &
      annotDf_df$`Molino-Surface_fish_AdjPVal` < 0.05 &
      (annotDf_df$`Pachón-Surface_fish_log2FoldChange` < -0.5 &
      annotDf_df$`Molino-Surface_fish_log2FoldChange` < -0.5), "red",
    "grey"), alpha = 0.6) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "black") +  # Add significance threshold
  geom_vline(xintercept = c(-0.5, 0.5), linetype = "dashed") + # Add foldchange threshold
  theme_light()+
  labs(x = "Log2 Fold Change", y = "-log10(P adjusted)", title = "Molino vs Surface fish") +
  theme(plot.title = element_text(hjust = 0.5)) 

vol_plot_mo_sf

#ggsave("output/vignette_run/median-norm/paper/SuppFig 2A Pachón vs Surface fish - common genes highlighted.png", plot = vol_plot_pa_sf, width = 6, height = 4, units = "in")
#ggsave("output/vignette_run/median-norm/paper/SuppFig 2A Pachón vs Surface fish - common genes highlighted.svg", plot = vol_plot_pa_sf, width = 6, height = 4, units = "in")
#ggsave("output/vignette_run/median-norm/paper/SuppFig 2A Molino vs Surface fish - common genes highlighted.png", plot = vol_plot_mo_sf, width = 6, height = 4, units = "in")
#ggsave("output/vignette_run/median-norm/paper/SuppFig 2A Molino vs Surface fish - common genes highlighted.svg", plot = vol_plot_mo_sf, width = 6, height = 4, units = "in")

rm(vol_plot_mo_sf,vol_plot_pa_sf)
```


## 4.3 Heat map

```{r}
# info
# For every KEGG pathways:
# When we copied the whole list from the website, we need to replace two spaces ("  ") to "" in the gene ID session.
```

### 4.3.1 TCA cycle

```{r}
# use TCA cycle as an example
# select specific genes in KEGG list
# https://www.genome.jp/entry/pathway+amex00020

tca <- read.csv("input/KEGG/Citrate cycle (TCA cycle) - Astyanax mexicanus (Mexican tetra).csv", header = F)


tca_test <- left_join(tca,annotDf_df,by = c("V1"="gene_id"))
tca_test <- na.omit(tca_test)

rownames(tca_test) <- tca_test$V1
tca_test<-tca_test[,11:19]

colnames(tca_test) <- gsub("Abundance..", "",colnames(tca_test))
colnames(tca_test) <- gsub("..Sample", "",colnames(tca_test))
tca_test <- log2(tca_test)

tca_test_1 <- expand.grid(gene.name = rownames(tca_test), individual_fish = colnames(tca_test))

tca_test_1$value[tca_test_1$individual_fish == "F8..129C..Molino"] <- tca_test$F8..129C..Molino
tca_test_1$value[tca_test_1$individual_fish == "F8..130N..Molino"] <- tca_test$F8..130N..Molino
tca_test_1$value[tca_test_1$individual_fish == "F8..130C..Molino"] <- tca_test$F8..130C..Molino
tca_test_1$value[tca_test_1$individual_fish == "F8..128N..Pachon"] <- tca_test$F8..128N..Pachon
tca_test_1$value[tca_test_1$individual_fish == "F8..128C..Pachon"] <- tca_test$F8..128C..Pachon
tca_test_1$value[tca_test_1$individual_fish == "F8..129N..Pachon"] <- tca_test$F8..129N..Pachon
tca_test_1$value[tca_test_1$individual_fish == "F8..126..Surface"] <- tca_test$F8..126..Surface
tca_test_1$value[tca_test_1$individual_fish == "F8..127N..Surface"] <- tca_test$F8..127N..Surface
tca_test_1$value[tca_test_1$individual_fish == "F8..127C..Surface"] <- tca_test$F8..127C..Surface

tca_test_z <- tca_test_1 %>%
  group_by(gene.name) %>%
  mutate(Z_Value = scale(value))


# Create a matrix for clustering
data_matrix <- dcast(tca_test_z, gene.name ~ individual_fish, value.var = "Z_Value")

# Perform hierarchical clustering on rows
row_dendrogram <- as.dendrogram(hclust(dist(data_matrix[, -1])))
order_rows <- order.dendrogram(row_dendrogram)

# Order the data frame according to the clustering
tca_test_z$gene.name <- factor(tca_test_z$gene.name, levels = data_matrix$gene.name[order_rows])


tca_test_heatmap <- ggplot(tca_test_z, aes(x = individual_fish, y = gene.name, fill = Z_Value)) +
  geom_tile() +
  scale_fill_gradient2(low = "darkblue", mid = "white", high = "darkred", midpoint = 0) +  # Set color gradient
  labs(title = "Heatmap for TCA cycle genes", x = "Morph", y = "Gene", fill = "Z Value") + # Add labels
  theme(plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 15),
        axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))


tca_test_heatmap

#ggsave("output/vignette_run/median-norm/median-norm-pathway-heatmaps/TCA cycles.png", plot = tca_test_heatmap, width = 12, height = 8, units = "in", dpi = 300)


rm(tca,tca_test,tca_test_1,tca_test_z,tca_test_heatmap,data_matrix,row_dendrogram,order_rows)

```

### 4.3.2 Ribosome

```{r}

# select specific genes in KEGG list
# https://www.genome.jp/entry/pathway+amex03050

ribosome <- read.csv("input/KEGG/Ribosome - Astyanax mexicanus (Mexican tetra) amex03010.csv", header = F)


ribosome_test <- left_join(ribosome,annotDf_df,by = c("V1"="gene_id"))
ribosome_test <- na.omit(ribosome_test)

# rownames(ribosome_test) <- ribosome_test$V1
# will generate errors
# Error in `.rowNamesDF<-`(x, value = value) : 
#  duplicate 'row.names' are not allowed
#In addition: Warning message:
#non-unique value when setting 'row.names': ‘103037375’ 
#ribosome_test<-ribosome_test[,11:19]

# Create unique row names
ribosome_test$RowName <- make.unique(as.character(ribosome_test$V1))

# Set the new unique names as row names
rownames(ribosome_test) <- ribosome_test$RowName
ribosome_test<-ribosome_test[,11:19]


colnames(ribosome_test) <- gsub("Abundance..", "",colnames(ribosome_test))
colnames(ribosome_test) <- gsub("..Sample", "",colnames(ribosome_test))
ribosome_test <- log2(ribosome_test)

ribosome_test_1 <- expand.grid(gene.name = rownames(ribosome_test), individual_fish = colnames(ribosome_test))

ribosome_test_1$value[ribosome_test_1$individual_fish == "F8..129C..Molino"] <- ribosome_test$F8..129C..Molino
ribosome_test_1$value[ribosome_test_1$individual_fish == "F8..130N..Molino"] <- ribosome_test$F8..130N..Molino
ribosome_test_1$value[ribosome_test_1$individual_fish == "F8..130C..Molino"] <- ribosome_test$F8..130C..Molino
ribosome_test_1$value[ribosome_test_1$individual_fish == "F8..128N..Pachon"] <- ribosome_test$F8..128N..Pachon
ribosome_test_1$value[ribosome_test_1$individual_fish == "F8..128C..Pachon"] <- ribosome_test$F8..128C..Pachon
ribosome_test_1$value[ribosome_test_1$individual_fish == "F8..129N..Pachon"] <- ribosome_test$F8..129N..Pachon
ribosome_test_1$value[ribosome_test_1$individual_fish == "F8..126..Surface"] <- ribosome_test$F8..126..Surface
ribosome_test_1$value[ribosome_test_1$individual_fish == "F8..127N..Surface"] <- ribosome_test$F8..127N..Surface
ribosome_test_1$value[ribosome_test_1$individual_fish == "F8..127C..Surface"] <- ribosome_test$F8..127C..Surface

ribosome_test_z <- ribosome_test_1 %>%
  group_by(gene.name) %>%
  mutate(Z_Value = scale(value))


# Create a matrix for clustering
data_matrix <- dcast(ribosome_test_z, gene.name ~ individual_fish, value.var = "Z_Value")

# Perform hierarchical clustering on rows
row_dendrogram <- as.dendrogram(hclust(dist(data_matrix[, -1])))
order_rows <- order.dendrogram(row_dendrogram)

# Order the data frame according to the clustering
ribosome_test_z$gene.name <- factor(ribosome_test_z$gene.name, levels = data_matrix$gene.name[order_rows])


ribosome_test_heatmap <- ggplot(ribosome_test_z, aes(x = individual_fish, y = gene.name, fill = Z_Value)) +
  geom_tile() +
  scale_fill_gradient2(low = "darkblue", mid = "white", high = "darkred", midpoint = 0) +  # Set color gradient
  labs(title = "Heatmap for ribosome genes", x = "Morph", y = "Gene", fill = "Z Value") + # Add labels
  theme(plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 15),
        axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))


ribosome_test_heatmap

#ggsave("output/vignette_run/median-norm/median-norm-pathway-heatmaps/ribosome.png", plot = ribosome_test_heatmap, width = 18, height = 12, units = "in", dpi = 300)


rm(ribosome,ribosome_test,ribosome_test_1,ribosome_test_z,ribosome_test_heatmap,data_matrix,row_dendrogram,order_rows)

```


### 4.3.3 Proteasome

```{r}

# select specific genes in KEGG list
# https://www.genome.jp/entry/pathway+amex03050

proteasome <- read.csv("input/KEGG/proteasome - Astyanax mexicanus (Mexican tetra).csv", header = F)


proteasome_test <- left_join(proteasome,annotDf_df,by = c("V1"="gene_id"))
proteasome_test <- na.omit(proteasome_test)

# rownames(proteasome_test) <- proteasome_test$V1
# will generate errors
# Error in `.rowNamesDF<-`(x, value = value) : 
#  duplicate 'row.names' are not allowed
#In addition: Warning message:
#non-unique value when setting 'row.names': ‘103037375’ 
#proteasome_test<-proteasome_test[,11:19]

# Create unique row names
proteasome_test$RowName <- make.unique(as.character(proteasome_test$V1))

# Set the new unique names as row names
rownames(proteasome_test) <- proteasome_test$RowName
proteasome_test<-proteasome_test[,11:19]


colnames(proteasome_test) <- gsub("Abundance..", "",colnames(proteasome_test))
colnames(proteasome_test) <- gsub("..Sample", "",colnames(proteasome_test))
proteasome_test <- log2(proteasome_test)

proteasome_test_1 <- expand.grid(gene.name = rownames(proteasome_test), individual_fish = colnames(proteasome_test))

proteasome_test_1$value[proteasome_test_1$individual_fish == "F8..129C..Molino"] <- proteasome_test$F8..129C..Molino
proteasome_test_1$value[proteasome_test_1$individual_fish == "F8..130N..Molino"] <- proteasome_test$F8..130N..Molino
proteasome_test_1$value[proteasome_test_1$individual_fish == "F8..130C..Molino"] <- proteasome_test$F8..130C..Molino
proteasome_test_1$value[proteasome_test_1$individual_fish == "F8..128N..Pachon"] <- proteasome_test$F8..128N..Pachon
proteasome_test_1$value[proteasome_test_1$individual_fish == "F8..128C..Pachon"] <- proteasome_test$F8..128C..Pachon
proteasome_test_1$value[proteasome_test_1$individual_fish == "F8..129N..Pachon"] <- proteasome_test$F8..129N..Pachon
proteasome_test_1$value[proteasome_test_1$individual_fish == "F8..126..Surface"] <- proteasome_test$F8..126..Surface
proteasome_test_1$value[proteasome_test_1$individual_fish == "F8..127N..Surface"] <- proteasome_test$F8..127N..Surface
proteasome_test_1$value[proteasome_test_1$individual_fish == "F8..127C..Surface"] <- proteasome_test$F8..127C..Surface

proteasome_test_z <- proteasome_test_1 %>%
  group_by(gene.name) %>%
  mutate(Z_Value = scale(value))


# Create a matrix for clustering
data_matrix <- dcast(proteasome_test_z, gene.name ~ individual_fish, value.var = "Z_Value")

# Perform hierarchical clustering on rows
row_dendrogram <- as.dendrogram(hclust(dist(data_matrix[, -1])))
order_rows <- order.dendrogram(row_dendrogram)

# Order the data frame according to the clustering
proteasome_test_z$gene.name <- factor(proteasome_test_z$gene.name, levels = data_matrix$gene.name[order_rows])


proteasome_test_heatmap <- ggplot(proteasome_test_z, aes(x = individual_fish, y = gene.name, fill = Z_Value)) +
  geom_tile() +
  scale_fill_gradient2(low = "darkblue", mid = "white", high = "darkred", midpoint = 0) +  # Set color gradient
  labs(title = "Heatmap for proteasome genes", x = "Morph", y = "Gene", fill = "Z Value") + # Add labels
  theme(plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 20),
        axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))


proteasome_test_heatmap

#ggsave("output/vignette_run/log2-norm-pathway-heatmaps/proteasome.png", plot = proteasome_test_heatmap, width = 12, height = 8, units = "in", dpi = 300)


rm(proteasome,proteasome_test,proteasome_test_1,proteasome_test_z,proteasome_test_heatmap,data_matrix,row_dendrogram,order_rows)

```




### Figure 2C candidate genes

```{r}
# select specific genes in KEGG list
# https://www.genome.jp/entry/pathway+amex00020

# read up and down regulated proteins in cavefish compared to surface fish
can_up <- read.csv("output/vignette_run/median-norm/DEG-analysis/annotated_upregulated in cavefish.csv")
can_down <- read.csv("output/vignette_run/median-norm/DEG-analysis/annotated_downregulated in cavefish.csv")

can <- rbind(can_up,can_down)

# only select genes I mention in the main text
can_test <- can[can$gene_name %in% c("decr1","acad11","gfpt1","tkta","tigarb","gmppb","cul3b","trim16","npepl1","pmpca","hint2","asrgl1","sarm1","agmat","slc12a9","npc1","jagn1b","rab10","rab1ab","timm8b","snx3","rpl5b","rpl18a","rpl31","rps12","rps16","rps27.1","eif4h","vars1","impdh1b"), ]

# rab10 twice
can_test[can_test$protein_id == "XP_007252241.1", "gene_name"] <- "rab10(1)"
can_test[can_test$protein_id == "XP_007260521.1", "gene_name"] <- "rab10(2)"

rownames(can_test) <- can_test$gene_name
can_test <- can_test[,12:20]

# Function to capitalize the first letter of a string (Protein name rules)
capitalize_first <- function(x) {
  substring(x, 1, 1) <- toupper(substring(x, 1, 1))
  x
}

# Apply the function to all row names
rownames(can_test) <- capitalize_first(rownames(can_test))

# change name to standard names
colnames(can_test) <- gsub("Abundance..", "",colnames(can_test))
colnames(can_test) <- gsub("..Sample", "",colnames(can_test))
colnames(can_test) <- gsub("F8..", "",colnames(can_test))

colnames(can_test) <- gsub("126..Surface", "Surface fish.1",colnames(can_test))
colnames(can_test) <- gsub("127N..Surface", "Surface fish.2",colnames(can_test))
colnames(can_test) <- gsub("127C..Surface", "Surface fish.3",colnames(can_test))
colnames(can_test) <- gsub("128N..Pachon", "Pachón.1",colnames(can_test))
colnames(can_test) <- gsub("128C..Pachon", "Pachón.2",colnames(can_test))
colnames(can_test) <- gsub("129N..Pachon", "Pachón.3",colnames(can_test))
colnames(can_test) <- gsub("129C..Molino", "Molino.1",colnames(can_test))
colnames(can_test) <- gsub("130N..Molino", "Molino.2",colnames(can_test))
colnames(can_test) <- gsub("130C..Molino", "Molino.3",colnames(can_test))

can_test <- log2(can_test)

can_test_1 <- expand.grid(gene.name = rownames(can_test), individual_fish = colnames(can_test))

can_test_1$value[can_test_1$individual_fish == "Molino.1"] <- can_test$Molino.1
can_test_1$value[can_test_1$individual_fish == "Molino.2"] <- can_test$Molino.2
can_test_1$value[can_test_1$individual_fish == "Molino.3"] <- can_test$Molino.3
can_test_1$value[can_test_1$individual_fish == "Pachón.1"] <- can_test$Pachón.1
can_test_1$value[can_test_1$individual_fish == "Pachón.2"] <- can_test$Pachón.2
can_test_1$value[can_test_1$individual_fish == "Pachón.3"] <- can_test$Pachón.3
can_test_1$value[can_test_1$individual_fish == "Surface fish.1"] <- can_test$`Surface fish.1`
can_test_1$value[can_test_1$individual_fish == "Surface fish.2"] <- can_test$`Surface fish.2`
can_test_1$value[can_test_1$individual_fish == "Surface fish.3"] <- can_test$`Surface fish.3`

can_test_z <- can_test_1 %>%
  group_by(gene.name) %>%
  mutate(Z_Value = scale(value))


# Create a matrix for clustering
data_matrix <- dcast(can_test_z, gene.name ~ individual_fish, value.var = "Z_Value")

# Perform hierarchical clustering on rows
#row_dendrogram <- as.dendrogram(hclust(dist(data_matrix[, -1])))
#order_rows <- order.dendrogram(row_dendrogram)

# Order the data frame according to the clustering
#can_test_z$gene.name <- factor(can_test_z$gene.name, levels = data_matrix$gene.name[order_rows])
can_test_z$individual_fish <- factor(can_test_z$individual_fish, levels = c("Surface fish.1","Surface fish.2","Surface fish.3","Pachón.1","Pachón.2","Pachón.3","Molino.1","Molino.2","Molino.3"))
can_test_z$gene.name <- factor(can_test_z$gene.name, levels = rev(c("Decr1","Acad11","Gfpt1","Tkta","Tigarb","Gmppb","Cul3b","Trim16","Npepl1","Pmpca","Hint2","Asrgl1","Sarm1","Agmat","Slc12a9","Npc1","Jagn1b","Rab10(1)","Rab10(2)","Rab1ab","Timm8b","Snx3","Rpl5b","Rpl18a","Rpl31","Rps12","Rps16","Rps27.1","Eif4h","Vars1","Impdh1b")))

can_test_heatmap <- ggplot(can_test_z, aes(x = individual_fish, y = gene.name, fill = Z_Value)) +
  geom_tile() +
  scale_fill_gradient2(low = "darkblue", mid = "white", high = "darkred", midpoint = 0) +  # Set color gradient
  theme_bw()+
  labs(title = "Heatmap of candidate genes", x = "Population", y = "Gene", fill = "Z Value") + # Add labels
  theme(plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 15),
        axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))


can_test_heatmap

#ggsave("output/vignette_run/median-norm/paper/Figure 2C heatmap of candidate genes.png", plot = can_test_heatmap, width = 5, height = 6, units = "in", dpi = 300)
#ggsave("output/vignette_run/median-norm/paper/Figure 2C heatmap of candidate genes.svg", plot = can_test_heatmap, width = 5, height = 6, units = "in", dpi = 300)

rm(can_up,can_down,can,can_test,can_test_1,can_test_z,can_test_heatmap,data_matrix,row_dendrogram,order_rows)
```

## 4.4 Bar plot for specific gene

### SuppFig 2B cbsa

```{r}

cbsa <- na.omit(annotDf_df[annotDf_df$gene_id == "103027202", 10:18])

cbsa_df <- data.frame(
  Population = rep(c("Molino", "Pachón", "Surface fish"), each = 3),
  Replicate = rep(c("Rep1", "Rep2", "Rep3"), times = 3),
  Value = as.numeric(as.character(cbsa))
)

cbsa_df$Population <- factor(cbsa_df$Population, levels = c("Surface fish", "Pachón", "Molino"))

box <- ggplot(cbsa_df, aes(x = Population, y = Value, fill = Population)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = pvg3) +
  geom_point(aes(x = Population), shape = 21, position = position_jitterdodge(dodge.width = 0.9)) +
  # Mannually add comparison and p value
  geom_segment(
    aes(x = 1, xend = 2, y = max(cbsa_df$Value) + 0.5, yend = max(cbsa_df$Value) + 0.5),
    color = "black", size = 1,
    inherit.aes = FALSE,
    data = data.frame(p.value = annotDf_df[annotDf_df$gene_id == "103027202", "Pachón-Surface_fish_AdjPVal"][1])) + # PA vs SF
  geom_segment(
    aes(x = 1, xend = 3, y = max(cbsa_df$Value) + 1, yend = max(cbsa_df$Value) + 1),
    color = "black", size = 1,
    inherit.aes = FALSE,
    data = data.frame(p.value = annotDf_df[annotDf_df$gene_id == "103027202", "Molino-Surface_fish_AdjPVal"][1])) + # MO vs SF
  
  annotate(
    "text",
    x = 1.5,
    y = max(cbsa_df$Value) + 0.6,
    label = paste("p = ", format.pval(annotDf_df[annotDf_df$gene_id == "103027202", "Pachón-Surface_fish_AdjPVal"][1], digits = 2)),
    hjust = 0.5, vjust = 0,
    size = 4, color = "black") + # PA vs SF
  annotate(
    "text",
    x = 2,
    y = max(cbsa_df$Value) + 1.1,
    label = paste("p = ", format.pval(annotDf_df[annotDf_df$gene_id == "103027202", "Molino-Surface_fish_AdjPVal"][1], digits = 2)),
    hjust = 0.5, vjust = 0,
    size = 4, color = "black") + # MO vs SF
  theme_minimal() +
  # HOW TO MAKE IT ITALIC
  #ylab(expression("" * italic("cbsa") * " protein normalized abundance")) +
  ylab(expression("Cbsa normalized protein abundance")) +
  labs(title = "cbsa protein abundance in 2-cell stage embryos") +
  theme(text = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = 0.5))  # Center the title

box

#ggsave("output/vignette_run/median-norm/paper/SuppFig 2B cbsa protein abundance in 2-cell stage embryos.png", plot = box, width = 6, height = 6, units = "in")
#ggsave("output/vignette_run/median-norm/paper/SuppFig 2B cbsa protein abundance in 2-cell stage embryos.svg", plot = box, width = 6, height = 6, units = "in")

rm(cbsa, cbsa_df, box)
```

### SuppFig 2B fastmyhc

```{r}

fastmyhc <- na.omit(annotDf_df[annotDf_df$gene_id == "103043447", 10:18])

fastmyhc_df <- data.frame(
  Population = rep(c("Molino", "Pachón", "Surface fish"), each = 3),
  Replicate = rep(c("Rep1", "Rep2", "Rep3"), times = 3),
  Value = as.numeric(as.character(fastmyhc))
)

fastmyhc_df$Population <- factor(fastmyhc_df$Population, levels = c("Surface fish", "Pachón", "Molino"))

box <- ggplot(fastmyhc_df, aes(x = Population, y = Value, fill = Population)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = pvg3) +
  geom_point(aes(x = Population), shape = 21, position = position_jitterdodge(dodge.width = 0.9)) +
  # Mannually add comparison and p value
  geom_segment(
    aes(x = 1, xend = 2, y = max(fastmyhc_df$Value) + 0.25, yend = max(fastmyhc_df$Value) + 0.25),
    color = "black", size = 1,
    inherit.aes = FALSE,
    data = data.frame(p.value = annotDf_df[annotDf_df$gene_id == "103043447", "Pachón-Surface_fish_AdjPVal"][3])) + # PA vs SF
  geom_segment(
    aes(x = 1, xend = 3, y = max(fastmyhc_df$Value) + 0.5, yend = max(fastmyhc_df$Value) + 0.5),
    color = "black", size = 1,
    inherit.aes = FALSE,
    data = data.frame(p.value = annotDf_df[annotDf_df$gene_id == "103043447", "Molino-Surface_fish_AdjPVal"][3])) + # MO vs SF
  
  annotate(
    "text",
    x = 1.5,
    y = max(fastmyhc_df$Value) + 0.3,
    label = paste("p = ", format.pval(annotDf_df[annotDf_df$gene_id == "103043447", "Pachón-Surface_fish_AdjPVal"][3], digits = 2)),
    hjust = 0.5, vjust = 0,
    size = 4, color = "black") + # PA vs SF
  annotate(
    "text",
    x = 2,
    y = max(fastmyhc_df$Value) + 0.55,
    label = paste("p = ", format.pval(annotDf_df[annotDf_df$gene_id == "103043447", "Molino-Surface_fish_AdjPVal"][3], digits = 2)),
    hjust = 0.5, vjust = 0,
    size = 4, color = "black") + # MO vs SF
  theme_minimal() +
  ylab(expression("Fast myhc normalized protein abundance")) +
  labs(title = "Fast myhc protein abundance in 2-cell stage embryos") +
  theme(text = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = 0.5))  # Center the title

box

#ggsave("output/vignette_run/median-norm/paper/SuppFig 2B fastmyhc protein abundance in 2-cell stage embryos.png", plot = box, width = 6, height = 6, units = "in")
#ggsave("output/vignette_run/median-norm/paper/SuppFig 2B fastmyhc protein abundance in 2-cell stage embryos.svg", plot = box, width = 6, height = 6, units = "in")

rm(fastmyhc, fastmyhc_df, box)
```

# SessionInfo

```{r}
sessionInfo()
```
